---
title: "Urban Fledgelings Analysis"
author: "Clint Malyk"
date: today
abstract-title: "Research question"
abstract: "How do environmental stresses and resource availability influence reproductive success?"
format: 
  html:
    embed-resources: true # get only a single html file and not an html in asupporting directory
    number-sections: true
editor: visual # opens qmd file in visual editor by default
editor_options: 
  chunk_output_type: console # output of code chunks goes to the console

# set document-wide code-chunk execution options:
execute:
  echo: true # true = print code-chunk in rendered html file
  warning: false # true = print warning messages in rendered html file after running code chunk
  error: false # true = print error messages in rendered html file after running code chunk
  message: false # true = print messages in rendered html file
---

## prepare the session, import and wrangle the data

```{r}
#| label: empty-environment
#| echo: false

rm(list= ls())
```

```{r}
#| label: import-libraries
#| echo: false

library(ggplot2) 
library(GGally) 
library(car) 
library(performance) 
library(tidyverse) 
library(pastecs)
library(DHARMa) 
library(ggeffects) # needed for predict_response
library(knitr) # -> to make nice tables
```

```{r}
#| label: import-data
#| echo: false

data <- read.csv("C:/Users/Clint/OneDrive - Universität Zürich UZH/Desktop/Ex7_folders/data/processed/clean.csv", stringsAsFactors = T, sep = ";") #-> converts "chr" to "factor" directly


```

## explore the data

### numerically

```{r}
#| label: tbl-descriptive-statistics
#| tbl-cap: "Descriptive statistics of data in data dataframe."
#| echo: false

knitr::kable(round(stat.desc(data, norm= T), 3)) # stat.desc creates descriptive stats 
```

In @tbl-descriptive-statistics we see that the P-value of the Shapiro-Wilk test for the outcome variable is \< 0.05, indicating that the outcome variable isn't normally distributed. Furthermore, the outcome is count data, so we use a poisson GLM to model the data.

### graphically

```{r}
#| label: fig-successes-histogram
#| fig-cap: "Distribution of number of fledglings observed."
#| echo: false

ggplot(data, aes(number_of_fledglings)) +
  geom_histogram(bins = 10) +
  labs(x = "Number of Fledglings")
```

```{r}
#| label: fig-fledglings-against-env
#| fig-cap: "Number of fledglings plotted against environmental noise in dB."
#| echo: false

ggplot(data, aes(environmental_noise_dB, number_of_fledglings)) +
  geom_point() +
  geom_smooth(method = "glm" , method.args = list(fam = poisson)) +
  labs(x = "Environmental Noise (dB)", y = "Number of Fledglings")
```

```{r}
#| label: fig-fledglings-against-resources
#| fig-cap: "Number of fledglings plotted against insect abundance in g/m\u00B2."
#| echo: false

ggplot(data, aes(Insect.Abundance_g.m2, number_of_fledglings)) +
  geom_point() +
  geom_smooth(method = "glm" , method.args = list(fam = poisson)) +
  labs(x = "Insect Abundance (g/m\u00B2)", y = "Number of Fledglings")
```

## fit candidate models & select preferred one

```{r}
#| label: candidate_models
#| echo: false

mod.00<- glm(number_of_fledglings ~ 1, data, family = poisson(link = log))
mod.01<- glm(number_of_fledglings~ environmental_noise_dB, data, family = poisson(link = log))
mod.02<- glm(number_of_fledglings~ Insect.Abundance_g.m2, data, family = poisson(link = log))
mod.03<- glm(number_of_fledglings~ environmental_noise_dB + Insect.Abundance_g.m2, data, family = poisson(link = log))
mod.04<- glm(number_of_fledglings~ environmental_noise_dB * Insect.Abundance_g.m2, data, family = poisson(link = log))

```

```{r}
#| label: compare_AICs

AIC(mod.00,mod.01,mod.02,mod.03,mod.04)

# select mod.03 & mod.04 based on AIC

```

```{r}
#| label: check_interaction_term

anova(mod.03, mod.04, test = "Chisq")

# the P-value of the Chisq test is > 0.05 -> interaction term not significant

```

## validate preferred model

```{r}
#| label: model_diagnostics
#| include: false

sim <- simulateResiduals(mod.03)
plot(sim)
```

```{r}
#| label: model_summary
#| output: false

summary(mod.03)

```

## interpret and report findings

Each 1 dB increase in noise corresponds to `r round(100 * abs(1 - exp(-0.00427)), 1)` % fewer fledglings (estimate = `r round(mod.03$coefficients[2],3)` , p = `r round(summary(mod.03)$coefficients[2,4],3)`). Each 1 g/m^2^ increase in insect biomass is associated with `r round(100 * abs(1 - exp(mod.03$coefficients[3])), 1)` % more fledglings (estimate = `r round(mod.03$coefficients[3],3)` , p = `r round(summary(mod.03)$coefficients[3,4],3)`).

## store session information

```{r}
#| label: session-info
#| include: false

version<- sessionInfo()

```
